### 2015-10-15 paper/images/diagram.tex
updates to paper

\documentclass[12pt]{article}
\usepackage{calc}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{lightgray}{gray}{0.9}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage[graphics, tightpage, active]{preview}
\usepackage{amsmath}

\setlength{\PreviewBorder}{2pt}
\PreviewEnvironment{tikzpicture}

\usetikzlibrary{shapes,arrows}
\usetikzlibrary{calc, positioning}
\usepackage{pgfplots}
\usepackage{filecontents}

%============================================================================================
%Additional commands
\tikzstyle{block} = [draw, rectangle, minimum height=2em, minimum width=4em]
%fill=blue!20
\tikzstyle{sum} = [draw, fill=blue!20, circle, node distance=1cm]
\tikzstyle{input} = [coordinate] \tikzstyle{output} = [coordinate]
\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]

%============================================================================================
\begin{document}
%============================================================================================

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------

 \def\radius{.7mm}
 \tikzstyle{branch}=[fill,shape=circle,minimum size=3pt]

\tikzstyle{dotted}= [dash pattern=on .01mm off 2.5mm,
                                         line cap=round]
 \begin{tikzpicture}[auto, node distance=2cm,>=latex']

    \node[block] (system) at (6, -2) {Server};
    \node[output] at (7.3, -2) (sys_output) {};
    \node[sum, left of=system] (sum)  at (4, -2) {};
    \node[block, above left = of sum, anchor=west, align = left] at (1,-1) (block1) {$p_1$};
    \node[block, left = of sum, anchor=west, align = left] at (1.6,-2) (block2) {$p_2$};
    \node[block, below left = of sum, anchor=west, align = left] at (1,-3) (block3) {$p_3$};

   \coordinate[below = 0.3cm of block1] (b1_turn);
   \coordinate[above = 0.3cm of block2] (b2_turn);
   \coordinate[above = 0.3cm of block3] (b3_turn);
   \coordinate(system_b1_turn) at (3.7, 1.55);
   \coordinate(system_b2_turn) at (3.7, -0.85);
   \coordinate(system_b3_turn) at (3.7, -3.265);

   \draw[->] (sum) -- node {$\bar x^{(t)}$} (system);
   \draw[-] (system) -- (sys_output);

   \draw[->] (block1.east) -| node [near start] {$\bar x_1^{(t)} $} (sum) ;
   \draw[->] (sys_output) |- (system_b1_turn) -| node [near start, above right = 0.0cm and 1.5cm of block1] {$(\bar \ell_{1}^{(t)}, \bar \ell^{(t-1)}_{1}, \dots)$} (block1.north);

   \draw[->] (block2) |- node [near end]  {$\bar x_2^{(t)} $} (sum);
   \draw[->] (sys_output) |- (system_b2_turn) -| node [near start, above right = 0.0cm and 1.5cm of block1] {$(\bar \ell_{2}^{(t)}, \bar \ell_2^{(t-1)}, \dots)$}  (block2.north);

   \draw[->] (block3) -| node [near start] {$\bar x_3^{(t)} $}  (sum);
   \draw[->] (sys_output) |- (system_b3_turn) -| node [near start, above right = 0.0cm and 1.5cm of block1] {$(\bar \ell_{3}^{(t)}, \bar \ell_3^{(t-1)}, \dots)$}  (block3.north);

\draw[dotted,line width = 2mm] (0.3,-5.1) -- (0.3, -5.8);

  \end{tikzpicture}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------

\end{document}